A Formal Model of Checked C

### 2022-08-07 sandp23/compilation.tex
update.

In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
In the compiled LLVM code, pointer access operations have different syntax when the modes are different.
\Cref{fig:flagtable} lists the different compiled syntax for a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang). The columns represent different pointer modes and the rows represent context modes.
For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation. In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$. if the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.

This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
with no loss of expressiveness;
as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
For the compiler formalism, 
we present a compilation algorithm that converts from
\lang to \elang, an untyped language without metadata
annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$.
The algorithm sheds
  light on how compilation can be implemented in the real Checked C
  compiler, while eschewing many important details (\elang has many 
  differences with LLVM IR).

@@

Compilation is defined by extending \lang's
typing judgment thusly:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each (NT-)array pointer variable to its mode and
each variable \code{p} to a pair of \emph{shadow
  variables} that keep \code{p}'s up-to-date upper and lower bounds; 
these may differ from the bounds in \code{p}'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound shadow variable is unnecessary; we include it for uniformity.}
% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
% complete judgment, implicitly assuming that $e$ is a well-typed and closed
% term.

@@

Due to space constraints, we explain the rules for compilation by
example, using a C-like syntax; the complete rules are given in
\iftr
Appendix~\ref{appx:comp1}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
and (c) generate right pointer accessing expressions based on modes.
%
A-normal form conversion is straightforward: compound expressions are
handled by storing results of subexpressions into temporary variables,
as in the following example.

@@

\myparagraph{Pointer Accesses and Modes}
%
In every declaration (or the beginning of a function body) of a pointer,
if the poniter is an (NT-)array one,
we first allocates two \emph{shadow variables}
to track the lower and upper bounds potentially changed for pointer arithmetic and NT-array bound widening.
We also associated with every $\cmode$-mode pointer variable according to its type;
we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
there is an inserted verification step similar to \Cref{fig:const-type},
which checks a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
In implementation, we actually optimize the verification away and substitute it
with the bounds and null-pointer checks.
Because tainted pointer is checked every time before it is used,
so that we only need to check the top pointer is 
well defined without recursively looking at the sub-terms in a nested pointer case.
The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
 and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
are computed based on the meet 
operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).

\myparagraph{Checked and Unchecked Blocks}
%
In the \systemname implementation,
$\euncheckedtext$ and $\echeckedtext$ blocks 
are compiled as context switching functions provided by the sandbox mechanism.
$\eunchecked{\overline{x}}{e}$ is compiled to 
$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
to execute expression $e$ with the arguments $\overline{x}$.
$\echecked{\overline{x}}{e}$ is compiled to 
$\texttt{callback}(\overline{x},e)$, where we perform 
a \texttt{callback} to a checked block code $e$ inside a sandbox.
In \systemname, We adopt an aggressive execution scheme that
directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
In the formalism, we rely on the type system to 
guarantee the context switching without creating the extra function calls for simplicity.

@@

\myparagraph{Function Pointers and Calls}
%
Function pointers are dealt with similarly to normal pointers,
but we insert checks to check if the pointer address is not null in 
the function store instead of heap, and whether or not the type is correctly represented, 
for both $\cmode$ and $\tmode$ mode pointers 
\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.

@@

We formalize both the compilation procedure and the simulation
theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
and then attempt to falsify it via Redex's support for random
testing. Redex allows us
  to specify compilation as logical rules (essentially, an extension
  of typing), but then execute it algorithmically to
  automatically test whether simulation holds. This process revealed
  several bugs in compilation and the theorem statement.
%
  % us gain confidence that our original pen and paper proof of
  % simulation remains true with the addition of variable bounds. }
We ultimately plan to prove simulation in the Coq model.

@@

Our random generator (discussed in the next section) never produces
$\euncheckedtext$ expressions (whose behavior could be undefined), so we
can only test a the simulation theorem 
as it applies to checked code. This limitation makes it
unnecessary to state the other direction of the simulation theorem
where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
that $e_0$ will never enter a stuck state if it is well-typed in
checked mode.

>>>>

In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions.
Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
In the compiled LLVM code, pointer access operations have different syntax when the modes are different.
\Cref{fig:flagtable} lists the different compiled syntax for a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang). The columns represent different pointer modes and the rows represent context modes.
For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation. In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$. If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.

This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
with no loss of expressiveness
as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
For the compiler formalism, 
we present a compilation algorithm that converts from
\lang to \elang, an untyped language without metadata
annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$.
The algorithm sheds
  light on how compilation can be implemented in the real Checked C
  compiler, while eschewing many vital details (\elang has many 
  differences with LLVM IR).

@@

Compilation is defined by extending \lang's
typing judgment as follows:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each (NT-)array pointer variable to its mode and
each variable \code{p} to a pair of \emph{shadow
  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
These may differ from the bounds in \code{p}'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound shadow variable is unnecessary; we include it for uniformity.}
% When $\Gamma$,$\Theta$ and $\rho$ are all empty, we write $e \gg \dot e$ rather than the
% complete judgment, implicitly assuming that $e$ is a well-typed and closed
% term.

@@

Due to space constraints, we explain the rules for compilation by
example, using a C-like syntax; the complete rules are given in
\iftr
Appendix~\ref{appx:comp1}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
and (c) generate right pointer accessing expressions based on modes.
%
A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
as in the following example.

@@

\myparagraph{Pointer Accesses and Modes}
%
In every declaration (or the beginning of a function body) of a pointer,
if the poniter is an (NT-)array one,
we first allocates two \emph{shadow variables}
to track the lower and upper bounds potentially changed for pointer arithmetic and NT-array bound widening.
Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
there is an inserted verification step similar to \Cref{fig:const-type},
which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
 and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
are computed based on the meet 
operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).

\myparagraph{Checked and Unchecked Blocks}
%
In the \systemname implementation,
$\euncheckedtext$ and $\echeckedtext$ blocks 
are compiled as context switching functions provided by the sandbox mechanism.
$\eunchecked{\overline{x}}{e}$ is compiled to 
$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
to execute expression $e$ with the arguments $\overline{x}$.
$\echecked{\overline{x}}{e}$ is compiled to 
$\texttt{callback}(\overline{x},e)$, where we perform 
a \texttt{callback} to a checked block code $e$ inside a sandbox.
In \systemname, we adopt an aggressive execution scheme that
directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
In the formalism, we rely on the type system to 
guarantee the context switching without creating the extra function calls for simplicity.

@@

\myparagraph{Function Pointers and Calls}
%
Function pointers are managed similarly to normal pointers,
but we insert checks to check if the pointer address is not null in 
the function store instead of heap, and whether or not the type is correctly represented, 
for both $\cmode$ and $\tmode$ mode pointers 
\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.

@@

We formalize both the compilation procedure and the simulation
theorem in the PLT Redex model we developed for \lang (see Sec.~\ref{sec:syntax}),
and then attempt to falsify it via Redex's support for random
testing. Redex allows us
  to specify compilation as logical rules (an extension
  of typing), but then execute it algorithmically to
  automatically test whether simulation holds. This process revealed
  several bugs in compilation and the theorem statement.
%
  % us gain confidence that our original pen and paper proof of
  % simulation remains true with the addition of variable bounds. }
We ultimately plan to prove simulation in the Coq model.

@@

As our random generator never generates
$\euncheckedtext$ expressions (whose behavior could be undefined), we can only test a the simulation theorem 
as it relates to checked code. This limitation makes it
unnecessary to state the other direction of the simulation theorem
where $e_0$ is stuck, because Theorem~\ref{thm:progress} guarantees
that $e_0$ will never enter a stuck state if it is well-typed in
checked mode.

### 2022-08-07 sandp23/formal.tex
update.

While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated; we can model mutable stack variables as pointers into the
mutable heap.
As mentioned, value $\evalue{0}{\tau}$
represents a $\enull$ pointer when $\tau$ is a pointer type;
correspondingly, $\heap(m,0)$ should always be undefined.
The relation steps to a \emph{result} $r$,
which is either an expression or a $\enull$ or $\ebounds$ failure,
representing a null-pointer dereference or out-of-bounds access,
respectively. Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterize undefined behavior.
%
The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the
stepped redex within $e$ was in a checked ($\cmode$) or
unchecked ($\umode$) region.

@@

\myparagraph{Checked and Tainted Pointer Operations}
%
The rules for pointer related operations---\textsc{S-DefC},
\textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
\textsc{S-DefNull}, and \textsc{S-Cast}.
The first five define deference and assignment operations---illustrate how the semantics checks bounds.
Rule \textsc{S-DefNull} transitions attempted null-pointer
dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a $\cmode$-mode
non-null (single) pointer.
When $\enull$ is returned by the
computation relation, the evaluation relation halts the entire
evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
dereference) is within the bounds designated by the pointer's annotation
and strictly less than the upper bound.

@@

\myparagraph{Binding and Function Calls}
%
The semantics handles variable scopes using the special $\erettext$
form. \textsc{S-Let} evaluates to a configuration whose expression is
$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
and remember $x$ and its the new value $\evalue{n}{\tau}$
in $e$'s scope that is defined by the $\erettext$ operation.
Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope;
after one-step evaluation is completed, 
we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
This procedure continues until $e'$ becomes a literal
$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
the literal.

@@

\myparagraph{Pointer Access}
%
Rules \textsc{T-Def} and \textsc{T-AssignArr} check array
dereference and assignment operations resp., returning the type of
pointed-to objects; rules for pointers for other object types are
similar.
The condition $m\le m'$ ensures that checked and unchecked pointers 
can only be deferenced in checked and unchecked regions, resp.;
The type rules do not attempt to reason whether the access is in bounds;
such check is deferred to the semantics.

@@

 The \textsc{T-CastPtr} rule
permits casting from an expression of type $\tau'$ to a checked pointer when
$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and it is built on the type equality
($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$); the many
rules ensure the relation is transitive. Most of the rules handle
casting between array pointer types. The second rule 
$0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
\tarrayptr{b_l}{b_h}{\tau}{m}$ permits treating a singleton
pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
There is another casting rule in \Cref{app:main} states that
 users are free to cast types in unchecked code regions, since unchecked regions can contain C code.

Since bounds expressions may
contain variables, determining assumptions like $b_l \leq b_l'$
requires reasoning about those variables' possible values. The type
system uses $\Theta$ to make such reasoning more precise.
$\Theta$ is a map from variables $x$ to
equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
It maps variables to equations that are recorded along the type checking procedure.
If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
\iftr
  Appendix~\ref{app:le}.
\else
  the supplemental report~\cite{checkedc-tech-report}.
\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.

@@

\myparagraph{Let Bindings and Dependent Function Pointers}
%
Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
type dependency. In particular, the result of evaluating a $\elettext$
may have a type that refers to one of its bound variables (e.g., if
the result is a checked pointer with a variable-defined bound); if so,
we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
Note that we restrict the expression $e_1$ to syntactically match the
structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).

Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
$\erettext$ does not appear in source programs but is introduced by the semantics when
evaluating a let binding (rule \textsc{S-Let} in
Fig.~\ref{fig:semantics}); this rule is needed for the preservation
proof.
% After the evaluation of a let binding a variable $x$ concludes,
%we need to restore any prior binding of $x$, which is either
%$\bot$ (meaning that there is no $x$ originally) or some value
%$\evalue{n}{\tau}$.

@@

In this subsection, we focus on our main meta-theoretic results about
\lang: type soundness (progress and preservation),
non-exposure, and non-crashing.
These proofs have been carried out in our Coq model.

@@

\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
A type environment $\Gamma$ is well-formed iff every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
\end{defi}

\begin{defi}[Heap Well-formedness]
For every $m$, A heap $\heap$ is well-formed iff (i) $\heap(m,0)$ is undefined, and
(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
contains no free variables. 
\end{defi}

\begin{defi}[Stack Well-formedness]
A stack snapshot $\varphi$ is well-formed iff
for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
contains no free variables. 
\end{defi}

@@

\begin{defi}[Stack Consistency]
A type environment $\Gamma$, variable predicate set $\Theta$, and
stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
\varphi$---iff for every variable $x$, $\Theta(x)$ is defined implies
$\Gamma(x) = \tau$ for some $\tau$ and 
$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
\end{defi}

\begin{defi}[Checked Stack-Heap Consistency]
A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---iff
for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
\end{defi}

\begin{defi}[Checked Heap-Heap Consistency]
A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---iff
for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
\end{defi}

>>>>

While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated. We can model mutable stack variables as pointers into the
mutable heap.
As mentioned, value $\evalue{0}{\tau}$
represents a $\enull$ pointer when $\tau$ is a pointer type.
Correspondingly, $\heap(m,0)$ should always be undefined.
The relation steps to a \emph{result} $r$,
which is either an expression or a $\enull$ or $\ebounds$ failure,
represent a null-pointer dereference or out-of-bounds access,
respectively. Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterizing undefined behavior.
%
The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the
stepped redex within $e$ was in a checked ($\cmode$) or
unchecked ($\umode$) region.

@@

\myparagraph{Checked and Tainted Pointer Operations}
%
The rules for pointer related operations---\textsc{S-DefC},
\textsc{S-DefT}, \textsc{S-AssignArrC}, \textsc{S-AssignArrT},
\textsc{S-DefNull}, and \textsc{S-Cast}.
The first five define deference and assignment operations and illustrate how the semantics checks bounds.
Rule \textsc{S-DefNull} transitions attempted null-pointer
dereferences to $\enull$, whereas \textsc{S-DefC} dereferences a $\cmode$-mode
non-null (single) pointer.
When $\enull$ is returned by the
computation relation, the evaluation relation halts the entire
evaluation with $\enull$ (using a rule not shown in Fig.~\ref{fig:c-context}); it
does likewise when $\ebounds$ is returned (see \Cref{sec:rem-semantics}).
\textsc{S-AssignArrC} assigns to an array as long as 0 (the point of
dereference) is within the bounds designated by the pointer's annotation
and strictly less than the upper bound.

@@

\myparagraph{Binding and Function Calls}
%
The semantics manages variable scopes using the special $\erettext$
form. \textsc{S-Let} evaluates to a configuration whose expression is
$\ret{x}{\evalue{n}{\tau}}{e})$. We keep $\varphi$ unchanged
and remember $x$ and its new value $\evalue{n}{\tau}$
in $e$'s scope that is defined by the $\erettext$ operation.
Every time when evaluation proceeds on $e$ (rule \textsc{S-RetCon}),
we install the stack value $\evalue{n}{\tau}$ for $x$ in $\varphi$ for the current scope.
After one-step evaluation is completed, 
we store $x$'s change in the result $\erettext$ operation $\ret{x}{\varphi'(x)}{e'})$,
and restore $x$'s outer score value $\varphi(x)$ in $\varphi'$. 
This procedure continues until $e'$ becomes a literal
$n\!:\!\tau$, in which case \textsc{S-RetEnd} removes the $\kw{ret}$ frame and returns
the literal.

@@

\myparagraph{Pointer Access}
%
\textsc{T-Def} and \textsc{T-AssignArr} rules examine array
dereference and assignment operations respectively, returning the type of
pointed-to objects. Rules for pointers for other object types are
similar.
The condition $m\le m'$ ensures that checked and unchecked pointers 
can only be dereferenced in checked and unchecked regions, respectively;
The type rules do not attempt to reason whether the access is in bounds;
such check is deferred to the semantics.

@@

 The \textsc{T-CastPtr} rule
permits casting from an expression of type $\tau'$ to a checked pointer when
$\tau' \sqsubseteq \tptr{\tau}{\cmode}$. This subtyping relation
$\sqsubseteq$ is given in Fig.~\ref{fig:checkc-subtype} and is built on the type equality
($\tau =_{\Theta} \tau'\Rightarrow\tau \sqsubseteq_{\Theta} \tau'$). The many
rules ensure the relation is transitive. Most of the rules manage
casting between array pointer types. The second rule 
$0\le b_l \wedge b_h \le 1 \Rightarrow \tptr{\tau}{m}\sqsubseteq
\tarrayptr{b_l}{b_h}{\tau}{m}$ permits treating a singleton
pointer as an array pointer with $b_h\le 1$ and $0 \le b_l$.
Two function pointer types are subtyped ($\tptr{\tfun{\overline{x}}{\overline{\tau}}{\tau}}{\xi} \sqsubseteq_{\Theta} \tptr{\tfun{\overline{x}}{\overline{\tau'}}{\tau'}}{\xi}$), 
if the output type are subtyped ($\tau\sqsubseteq_{\Theta}\tau'$) and the argument types are reversely subtyped ($\overline{\tau'}\sqsubseteq_{\Theta}\overline{\tau}$).
There is another casting rule in \Cref{app:main} stating that
 users are free to cast types in unchecked code regions, since unchecked regions can contain C code.

Since bounds expressions may
contain variables, determining assumptions like $b_l \leq b_l'$
requires reasoning about the probable values of these variables'. The type
system uses $\Theta$ to make such reasoning more precise.
$\Theta$ is a map from variables $x$ to
equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
It maps variables to equations that are recorded along the type checking procedure.
If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
\iftr
  Appendix~\ref{app:le}.
\else
  the supplemental report~\cite{checkedc-tech-report}.
\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.

@@

\myparagraph{Let Bindings and Dependent Function Pointers}
%
Rules \textsc{T-Let} and \textsc{T-LetInt} type a $\elettext$ expression, which also admits
type dependency. In particular, the result of evaluating a $\elettext$
may have a type that refers to one of its bound variables (e.g., if
the result is a checked pointer with a variable-defined bound). If so, we must substitute away this variable once it goes out of scope (\textsc{T-LetInt}). 
Note that we restrict the expression $e_1$ to syntactically match the
structure of a Bounds expression $b$ (see Fig.~\ref{fig:checkc-syn}).

Rule \textsc{T-RetInt} types a $\erettext$ expression when $x$ is of type $\tint$.
$\erettext$ does not appear in source programs but is introduced by the semantics when
evaluating a let binding (rule \textsc{S-Let} in
Fig.~\ref{fig:semantics}). This rule is needed for the preservation
proof.
% After the evaluation of a let binding a variable $x$ concludes,
%we need to restore any prior binding of $x$, which is either
%$\bot$ (meaning that there is no $x$ originally) or some value
%$\evalue{n}{\tau}$.

@@

In this subsection, we focus on our main meta-theoretic results about
\lang: type soundness (progress and preservation),
non-exposure, and non-crashing.
These proofs have been conducted in our Coq model.

@@

\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
\end{defi}

\begin{defi}[Heap Well-formedness]
For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
contains no free variables. 
\end{defi}

\begin{defi}[Stack Well-formedness]
A stack snapshot $\varphi$ is well-formed if
for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
contains no free variables. 
\end{defi}

@@

\begin{defi}[Stack Consistency]
A type environment $\Gamma$, variable predicate set $\Theta$, and
stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
$\Gamma(x) = \tau$ for some $\tau$ and 
$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
\end{defi}

\begin{defi}[Checked Stack-Heap Consistency]
A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
\end{defi}

\begin{defi}[Checked Heap-Heap Consistency]
A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
\end{defi}
